---
title: "Reformatting"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Remapping}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dm)
library(dunlin)
```

## Introduction

Reformatting in `dulin` consists in replacing predetermined values by another in particular variables for selected
tables of a data set stored under the form of a `dm` object.

This is performed in two steps: 

1. A Reformatting Map (`list` object) is created which specifies the correspondence between the old and the new values

2. The reformatting itself is performed with the `dunlin::apply_reformat()` function.

## The Formatting Map Structure

The Reformatting Map is a list object with 4 hierarchic levels. 

1. The Table Name: Typically one of the values returned by `names(db)` where db is `dm` object. For instance, in the case of an `adam` data set, `adsl` or `adae` etc...

2. The Variable Name: The name of the column in a table of a `dm` object. For instance `ACTARM`.

3. The Key: the new value that will replace the old ones

4. The Values: a vector or values, that have to be replaced by their associated key.
 
The `All` keyword, in HIGHER, lower or Mixed case, can be used instead of a table name to indicate that a particular
variable should be changed in every table where it appears.
 
The values can be a `string`, a `character` vector or `NA`. Values that are present in the original table and not listed
in the mapping are left unchanged, except that all altered columns are turned into `factor` class. If the original
variable is a `factor`, the non-transformed levels are preserved even if the corresponding value does not appear.

The order in which the keys appear determine the order of the levels in the output. Hence the function can be used to re
level variables. However, the level replacing missing values will automatically be placed last.
 
#### Example of Reformatting Map

```{r}

my_map <- list(
  airlines = list( # This is the Table Name.
    name = list( # This is the Variable Name.
      "AE" = c("American Airlines Inc."), # AE is the Key, `c("American Airlines Inc.")` is the value.
      "Alaska and Hawaiian Airlines" = c("Alaska Airlines Inc.", "Hawaiian Airlines Inc.")
    )
  ),
  planes = list(
    manufacturer = list(
      "Airbus" = "AIRBUS INDUSTRIE",
      "New Level" = "new_level",
      "<Missing>" = NA
    ),
    model = list(
      "EMB-145" = c("EMB-145XR"),
      "Other 737" = c("737-824", "737-724", "737-732")
    )
  ),
  All = list(
    year = list(
      "Current" = "2013"
    )
  )
)
```
 

## Re Formatting
 
Once the re formatting map is defined, the reformatting can be performed on a `dm` object.

#### Example

```{r}

db <- dm::dm_nycflights13()

db <-
  db %>%
  dm_zoom_to(planes) %>%
  mutate(manufacturer = ifelse(manufacturer == "EMBRAER", NA, manufacturer)) %>%
  dm_update_zoomed()

res <- apply_reformat(db, format = my_map)
head(db$airlines$name)
head(res$airlines$name)

# New levels are introduced and missing values can be replaced.
head(db$planes$manufacturer)
head(res$planes$manufacturer)

# Using the `All` keyword changes several tables.
head(db$weather$year)
head(res$weather$year)

head(db$flights$year)
head(res$flights$year)
```

## Reformatting Map with `NULL` values

Reformatting functions tolerate `NULL` values in the reformatting maps. In the following example, the table `df2` will
be left unchanged as well as the `logi` column of the `df1` table. In the `char` column, only `a` and `k` will be
reformatted and the column changed to `factor`. `B` wont be included in the levels.

```{r}

# This is a valid reformatting map
null_map <- list(
    df1 = list(
      char = list(
        "A" = c("a", "k"),
        "B" = NULL
      ),
      logi = NULL
    ),
    df2 = NULL
  )

assert_reformat(null_map)

```
